---
title: "Homework 1"
author: "Amin Yakubu"
date: "2/23/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
```

```{r}
library(tidyverse)
library(glmnet)
library(caret)
library(corrplot)
library(plotmo)
library(pls)
```

# Data 

```{r}
train = read_csv('./data/solubility_train.csv')
test = read_csv('./data/solubility_test.csv')
```

The data has been divided into training and testing. The testing data has 951 observations and the test data has 356 observations. There are `r ncol(train)` predictors. `r sum(apply(train,2,function(x) { all(x %in% 0:1) }))` are binary variables that indicate the presence or absence of a particular chemical substructure, 16 are count descriptors, such as the number of bonds or the number of bromine atoms, and 4 are continuous descriptors, such as molecular weight or surface area. The response is in the column `Solubility` which is a continuous variable


```{r}
# Checking for missing values
missing_train <- sapply(train, function(x) sum(is.na(x)))
missing_train[missing_train > 0]

missing_test <- sapply(test, function(x) sum(is.na(x)))
missing_test[missing_test > 0]
```

No missing data 

```{r}
plot(train$Solubility)
```

# Linear Model

```{r}
lm.fit = lm(Solubility ~ ., data = train)
summary(lm.fit)
```

# Mean square error using the test data

```{r}
lm.pred = predict(lm.fit, newdata = test[,-ncol(test)])

mean((lm.pred - test$Solubility)^2)
```

The mean square error using the test data is 0.55589. 

Data processing 
```{r}
#Training set
X.train = model.matrix(Solubility ~ ., train)[,-1]
y.train = train$Solubility

#Testing set
X.test = model.matrix(Solubility ~ ., test)[,-1]
y.test = test$Solubility

```

Fit a ridge regression model on the training data, with λ chosen by cross-validation. Report the test error

```{r}
ctrl1 <- trainControl(method = "repeatedcv", number = 10, repeats = 5)

set.seed(1)
ridge.fit <- train(X.train, y.train,
                     method = "glmnet",
                     tuneGrid = expand.grid(alpha = 0, 
                                            lambda = exp(seq(-10,10, length = 200))),
                     trControl = ctrl1)

plot(ridge.fit)
plot(ridge.fit, xTrans = function(x) log(x)) # here were are plotting log lambda so it looks like the previous plots

ridge.fit$bestTune

```

```{r}
bestlam.ridge = ridge.fit$lambda.min
bestlam.ridge
```

```{r}
ridge.pred = predict(ridge.fit$finalModel, s = bestlam.ridge, newx = X.test)
mean((ridge.pred - y.test)^2)
```

The mean test error is 0.51346

# The Lasso

```{r}
ctrl1 <- trainControl(method = "repeatedcv", number = 10, repeats = 5)

set.seed(1)
lasso.fit <- train(X.train, y.train,
                     method = "glmnet",
                     tuneGrid = expand.grid(alpha = 1, 
                                            lambda = exp(seq(-10,10, length = 200))),
                     trControl = ctrl1)

plot(lasso.fit)
plot(lasso.fit, xTrans = function(x) log(x)) # here were are plotting log lambda so it looks like the previous plots

names(lasso.fit)
lasso.fit$bestTune
```

```{r}
bestlam.lasso = lasso.fit$bestTune$lambda
bestlam.lasso
```

```{r}
lasso.pred = predict(lasso.fit$finalModel, s = bestlam.lasso, newx = X.test)
mean((lasso.pred - y.test)^2)
```

The mean error is 0.4987. 

```{r}
lasso.coef = predict(lasso.fit$finalModel, type = "coefficients", s = bestlam.lasso)[1:ncol(train),]
length(lasso.coef)
```

```{r}
length(lasso.coef[lasso.coef != 0])
```

There are 144 non-zero coefficient 

# PCR 

```{r}
ctrl1 <- trainControl(method = "repeatedcv", number = 10, repeats = 5)

set.seed(2)
pcr.fit <- train(X.train, y.train,
                  method = "pcr",
                  tuneLength = 228,
                  trControl = ctrl1,
                  scale = TRUE)

pred.pcr <- predict(pcr.fit$finalModel, newdata = X.test, 
                       ncomp = pcr.fit$bestTune$ncomp)

mean((pred.pcr - y.test)^2)

ggplot(pcr.fit, highlight = TRUE) + theme_bw()

```



